---
title: "The Statistics of Early Childhood Education in Indonesia"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
       
---


```{r setup, include=FALSE}
pacman::p_load(rio, here, tidyverse, maps, mapdata, ggthemes, ggrepel, rgdal,raster, viridis,sf,terra, colorspace, flexdashboard, plotly,janitor, maptools,ggthemes, viridis)
```



# Overview 

Sidebar Title {.sidebar}
--------------

**General information about ECE Programs in Indonesia**

The First Page of this dashboard contains overall nationwide data about Total number of ECE Programs, total number of students enrolled in all ECE programs, total numbers of ECE Teachers and Facilitators, and total number of classrooms. 

*Data source:*
Data is taken from the 2020/2021 Annual ECE Statistics Report issued by the Ministry of Education and Cultural. This data reflects ECE Programs statistics in Indonesia as off 30 November 2020.


-------------

```{r include=FALSE}
#Load Province name
name <- import(here("data", "name.xlsx"))
name$id <- as.character(name$id)

#Load Indonesia shapefile
idn_shape2 <- rgdal::readOGR(dsn=path.expand("/Users/merlyklaas/Documents/University of Oregon/2022-2023/EDLD 655 Capstone/capstone-project/data/smallest/idn_admbnda_adm1_bps_20200401.shp"),layer = "idn_admbnda_adm1_bps_20200401")
st_drivers()
idn_shape_df <- fortify(idn_shape2)
head(idn_shape_df)
idn <- left_join(idn_shape_df, name)

#Load ECE Data
stat <- import(here("data","stat.xlsx")) %>% 
  clean_names() %>%
  mutate(across(setdiff(names(select(., where(is.character))), 'province'), as.numeric))
dat <- left_join(idn, stat)

str(stat)
```

Column {.tabset data-width=700}
-----------------------------------------

### Number of ECE Programs



```{r include=FALSE}
totece <- dat %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(x=long,y=lat, group=group, fill=institution, text=province), color='black')+
  scale_fill_continuous_divergingx(palette = "Earth")+
  ggthemes::theme_map(base_size=8)+
  theme(strip.background = element_blank(),
        legend.position ="bottom",
        legend.direction = "horizontal",
        legend.title = element_text(size = 6), 
        legend.text  = element_text(size = 6),
        legend.key.size = unit(1, "lines"),
        plot.title.position = "plot",
        text=element_text(family= "sans"),
        plot.title = element_text(face="bold"),
        plot.subtitle = element_text(face="italic"),
        plot.caption = element_text(color ="gray30",hjust = 0),
        axis.title = element_text(face = "bold"))+
  labs(fill = "# of ECE Programs",
       title = "ECE Programs Availability by Province in Indonesia",
       subtitle = "Java islands have the highest number of ECE Programs",
       x="",
       y="")
totece
```

```{r}
ggplotly(totece)
```


### Students Enrollment

```{r include=FALSE}
totstu<- dat %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(x=long,y=lat, group=group, fill=participants, text=province), color='black')+
  scale_fill_continuous_divergingx(palette = "Earth")+
  ggthemes::theme_map(base_size=8)+
  theme(strip.background = element_blank(),
        legend.position ="bottom",
        legend.direction = "horizontal",
        legend.title = element_text(size = 6), 
        legend.text  = element_text(size = 6),
        legend.key.size = unit(1, "lines"),
        plot.title.position = "plot",
        text=element_text(family= "sans"),
        plot.title = element_text(face="bold"),
        plot.subtitle = element_text(face="italic"),
        plot.caption = element_text(color ="gray30",hjust = 0),
        axis.title = element_text(face = "bold"))+
  labs(fill = "# of Students Enrolled",
       title = "Early Childhood Students Enrollment in ECE Program  by Province in Indonesia",
       subtitle = "Java island has the highest students enrollment in ECE Programs",
       x="",
       y="")
```
```{r}
ggplotly(totstu)
```

### Teacher & Facilitator
```{r include=FALSE}
totteach<- dat %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(x=long,y=lat, group=group, fill=teacher_facilitator, text=province), color='black')+
  scale_fill_continuous_divergingx(palette = "Earth")+
  ggthemes::theme_map(base_size=8)+
  theme(strip.background = element_blank(),
        legend.position ="bottom",
        legend.direction = "horizontal",
        legend.title = element_text(size = 6), 
        legend.text  = element_text(size = 6),
        legend.key.size = unit(1, "lines"),
        plot.title.position = "plot",
        text=element_text(family= "sans"),
        plot.title = element_text(face="bold"),
        plot.subtitle = element_text(face="italic"),
        plot.caption = element_text(color ="gray30",hjust = 0),
        axis.title = element_text(face = "bold"))+
  labs(fill = "# of Teacher",
       title = "Teachers and Facilitors in ECE Program  by Province in Indonesia",
       subtitle = "Java island has the highest students enrollment in ECE Programs",
       x="",
       y="")

```

```{r}
ggplotly(totteach)
```

# Student to Teacher Ratio

Sidebar Title {.sidebar}
--------------

**Student per Teacher Ratio**

The Second Page of this dashboard contains student to teacher ratio by type of ECE Programs aggregated in Province level. There are two main types of ECE Program, Non-formal ECE and Formal ECE. 

There are three types of Non-Formal ECE:

+ Daycare *(Tempat Penitipan Anak)* for	0-6 years old with the Priority for 0-4 years old	
+ Playgroup *(Kelompok Bermain)*	for 2-6 years old with Priority for 3-4 years old	
+ Integrated ECE *(Satuan Paud Sejenis)* for	0-6 years old

And only one type of Formal ECE
+ Kindergarten (Taman Kanak-kanak/ Raudatul Athfal)	for 4-6 years old with priority for 5-6 years old


Thus, in Permendikbud 137 year 2014 about National ECE Standards, the government regulates the maximum student per teacher ratio based on children age:

+ aged 0-2 = 1:4
+ aged 2-4 = 1:8
+ aged 4-6 = 1:15



*Data source:*
Data is taken from the 2020/2021 Annual ECE Statistics Report issued by the Ministry of Education and Cultural. This data reflects ECE Programs statistics in Indonesia as off 30 November 2020.

-------------


Column {.tabset data-width=700}
-----------------------------------------

### Kindergarten

```{r}

tkratio<- dat %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(x=long,y=lat, group=group, fill=tk_stu_teacher, text=province), color='black')+
   scale_fill_viridis_c(option = "magma", direction=-1)+
  #scale_fill_continuous_divergingx(palette = "Geyser")+
  ggthemes::theme_map(base_size=8)+
  theme(strip.background = element_blank(),
        legend.position ="bottom",
        legend.direction = "horizontal",
        legend.title = element_text(size = 6), 
        legend.text  = element_text(size = 6),
        legend.key.size = unit(1, "lines"),
        plot.title.position = "plot",
        text=element_text(family= "sans"),
        plot.title = element_text(face="bold"),
        plot.subtitle = element_text(face="italic"),
        plot.caption = element_text(color ="gray30",hjust = 0),
        axis.title = element_text(face = "bold"))+
  labs(fill = "Ratio",
       title = "Student to Teacher Ratio in Kindergarten",
    #   subtitle = "Java island has the highest students enrollment in ECE Programs",
       x="",
       y="")
tkratio
```

```{r}

```

### Daycare

```{r}
tparatio<- dat %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(x=long,y=lat, group=group, fill=tpa_stu_teach, text=province), color='black')+
   scale_fill_viridis_c(option = "magma", direction=-1)+
  #scale_fill_continuous_divergingx(palette = "Geyser")+
  ggthemes::theme_map(base_size=8)+
  theme(strip.background = element_blank(),
        legend.position ="bottom",
        legend.direction = "horizontal",
        legend.title = element_text(size = 6), 
        legend.text  = element_text(size = 6),
        legend.key.size = unit(1, "lines"),
        plot.title.position = "plot",
        text=element_text(family= "sans"),
        plot.title = element_text(face="bold"),
        plot.subtitle = element_text(face="italic"),
        plot.caption = element_text(color ="gray30",hjust = 0),
        axis.title = element_text(face = "bold"))+
  labs(fill = "Ratio",
       title = "Student to Teacher Ratio in Daycare",
   #    subtitle = "Java island has the highest students enrollment in ECE Programs",
       x="",
       y="")
```
```{r}
ggplotly(tparatio)
```

### Playgroup

```{r}
kbratio<- dat %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(x=long,y=lat, group=group, fill=kb_stu_teach, text=province), color='black')+
   scale_fill_viridis_c(option = "magma", direction=-1)+
  #scale_fill_continuous_divergingx(palette = "Geyser")+
  ggthemes::theme_map(base_size=8)+
  theme(strip.background = element_blank(),
        legend.position ="bottom",
        legend.direction = "horizontal",
        legend.title = element_text(size = 6), 
        legend.text  = element_text(size = 6),
        legend.key.size = unit(1, "lines"),
        plot.title.position = "plot",
        text=element_text(family= "sans"),
        plot.title = element_text(face="bold"),
        plot.subtitle = element_text(face="italic"),
        plot.caption = element_text(color ="gray30",hjust = 0),
        axis.title = element_text(face = "bold"))+
  labs(fill = "Ratio",
       title = "Student to Teacher Ratio in Playgroup",
   #    subtitle = "Java island has the highest students enrollment in ECE Programs",
       x="",
       y="")
```
```{r}
ggplotly(kbratio)
```

### Integrated ECE

```{r}
spsratio<- dat %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(x=long,y=lat, group=group, fill=sps_stu_teach, text=province), color='black')+
   scale_fill_viridis_c(option = "magma", direction=-1)+
  #scale_fill_continuous_divergingx(palette = "Geyser")+
  ggthemes::theme_map(base_size=8)+
  theme(strip.background = element_blank(),
        legend.position ="bottom",
        legend.direction = "horizontal",
        legend.title = element_text(size = 6), 
        legend.text  = element_text(size = 6),
        legend.key.size = unit(1, "lines"),
        plot.title.position = "plot",
        text=element_text(family= "sans"),
        plot.title = element_text(face="bold"),
        plot.subtitle = element_text(face="italic"),
        plot.caption = element_text(color ="gray30",hjust = 0),
        axis.title = element_text(face = "bold"))+
  labs(fill = "Ratio",
       title = "Student to Teacher Ratio in Integrated ECE",
   #    subtitle = "Java island has the highest students enrollment in ECE Programs",
       x="",
       y="")
```
```{r}
ggplotly(spsratio)
```


